Optimal forward modeling of gravitational fields using tesseroids with adaptive discretization

### 2014-09-01 body.tex
Copied shell equations to text

The gravitational potential,
gravitational attraction,
and
Marussi (gravity gradient) tensor
caused by the homogeneous spherical half-shell
at a computation point
$P = (\lambda=90^\circ, \phi=0^\circ, r=r_2+h)$
are

\begin{equation}
    V(h) = 2\pi G \rho \left[
        \frac{l^3 + {r'}^3}{3(r_2 + h)} - 0.5 {r'}^2 \right]
         \Biggr|_{r'=r_1}^{r'=r_2},
    \label{eq:shellpot}
\end{equation}
\begin{equation}
    g_z(h) = 2\pi G \rho \left[
        \frac{l^3 + {r'}^3}{3(r_2 + h)^2} - l \right]
        \Biggr|_{r'=r_1}^{r'=r_2},
    \label{eq:shellgz}
\end{equation}
\begin{equation}
    g_{xx}(h) = g_{yy}(h) = -\frac{1}{2} g_{zz}(h),
\end{equation}
\begin{equation}
    g_{xy}(h) = g_{xz}(h) = g_{yz}(h) = 0,
\end{equation}

@@

\begin{equation}
    g_{zz}(h) = 2\pi G \rho \left[
        2\frac{l^3 + {r'}^3}{3(r_2 + h)^3}
        - \frac{l}{r_2 + h} + \frac{r_2 + h}{l}
        \right]
        \Biggr|_{r'=r_1}^{r'=r_2},
    \label{eq:shellgzz}
\end{equation}

\noindent
where $\rho$ is the density,
$l = \sqrt{(r_2 + h)^2 + {r'}^2}$,
and
$r_1$ and $r_2$ are radial coordinates of
the bottom and top of the spherical shell,
respectively.

For our comparison,
we used a spherical half-shell with
$\rho=2.7\ g.cm^{-3}$,
$r_1=6378.137\ km$ (mean Earth radius),
and
$r_2 = r_1 + 50\ km$.
We discretized the half-shell
into tesseroids of equal size
and calculated the gravitational effects of
the half-shell
(using equations \ref{eq:shellpot}-\ref{eq:shellgzz})
and the corresponding tesseroid model
at various heights.
These calculations were repeated
for varying values of
the distance/size ratio $R$.
This analysis
allows us to establish
the optimal value of $R$
that achieves a desired error level.

Data and source code to produce the results of this comparison
are provided by \textbf{(cite notebook on figshare)}.

@@

The methods described above
were implemented the software package
Tesseroids version 1.1.1.
The package consists
of a range of command-line programs
written in the C programming language.
It is freely available
online\footnote{www.leouieda.com/tesseroids}
under the terms of
the open-source BSD 3-clause license.
\textbf{(cite archive on figshare)} provides a persistent
archive for the source code and compiled binaries of Tesseroids.

@@

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam eu dolor pretium,
egestas mauris sed, dapibus quam. Duis hendrerit mollis nunc a consequat. Nulla
et sem consectetur, interdum velit eget, aliquam ipsum. Praesent sagittis
tortor diam, sed ultrices magna ullamcorper vitae. Proin vitae orci augue.
Morbi dictum ligula gravida sem malesuada facilisis. Mauris nibh metus, cursus
eget imperdiet vitae, pretium at lorem. Praesent nisi mauris, pretium ut risus
fermentum, egestas tincidunt nibh. Mauris nulla orci, consequat eu pharetra
non, mattis ut urna. Mauris facilisis orci eros. Nam mattis non magna iaculis
consectetur. Morbi sodales dolor vitae felis sagittis, eget faucibus turpis
convallis. Nullam malesuada, mauris et ultricies rutrum, odio nulla gravida
nunc, ac volutpat eros lectus eget lacus. Integer venenatis velit vel justo
pellentesque, quis molestie sem vestibulum.

>>>>

The gravitational potential,
gravitational attraction,
and
Marussi (gravity gradient) tensor
of a homogeneous spherical shell
are \citep{Grombein2013}

\begin{equation}
V(r) = \frac{4}{3}\pi G \rho \frac{r_2^3 - r_1^3}{r},
\end{equation}

\begin{equation}
g_z(r) = \frac{V(r)}{r},
\end{equation}

\begin{equation}
g_{zz}(r) = 2\frac{V(r)}{r^2},
\end{equation}

\begin{equation}
g_{xx}(r) = g_{yy}(r) = -\frac{g_{zz}}{2},
\end{equation}

@@

\begin{equation}
g_x(r) = g_y(r) = g_{xy}(r) = g_{xz}(r) = g_{yz}(r) = 0,
\end{equation}

\noindent
where $r$ is the radial coordinate of the computation point and
$r_1$ and $r_2$ are radial coordinates
of the bottom and top of the spherical shell, respectively.

@@

The recursive division algorithm described above
is implemented in the software package
Tesseroids.
The package consists
of command-line programs
written in the C programming language.
Tesseroids is open-source software
and is freely available
on the internet\footnote{www.leouieda.com/tesseroids}
\textbf{CITE ARCHIVE} provides a persistent
archive for the source code.

The code used to produce the results and figures presented here
is available at \textbf{DOI LINK AND CITATION}.

@@
